---
title: "IUCN Tutorial"
author: "Cerren Richards"
date: "18/09/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Importing species' threat data from the IUCN

To identify which species are threatened from oil drilling, shipping lanes and subsistence fishing, we first must extract all the seabird threats from the IUCN Threat Classification Scheme: https://www.iucnredlist.org/resources/threat-classification-scheme. 

Note - Because threats are updated through time, they may change in the future.

Note - an API key is needed to download the IUCN data. APIs can be requested from: https://apiv3.iucnredlist.org/api/v3/token

Here, for an example, we use the list of seabird downloaded from Richards et al. 2021:  https://datadryad.org/stash/dataset/doi%253A10.5061%252Fdryad.x69p8czhd 


```{r message=FALSE, error=FALSE, warning=FALSE, eval = FALSE}

# Load packages
library(rredlist); library(rlist) 

# Load the names of seabirds
seabirds <- read.csv("Imputed Trait Data.csv")

# Create a vector of species names from the traits dataframe
# IMPORTANT - Species names must match the IUCN species names
spp <- as.character(seabirds$binomial) 

# Enter your unique IUCN code here
# Request from: https://apiv3.iucnredlist.org/api/v3/token 
iucn_key <- "" 

# Extract the threats for each species from th IUCN database.
# Creates a list
# NOTE - Depending on the size of your species list, this can take a while
# For 341 species, it takes 10-15 minutes to run
iucn <- lapply(spp, function(x) {
  y <- rl_threats(name = x, key = iucn_key)
  Sys.sleep(2)
  # 2 second delay makes API work better - recommended by IUCN
  return(y)
})

### Add an extra column to the dataframes in the list with the binomial names:
for (i in 1:length(iucn)) { 
  iucn[[i]][["result"]]$binomial <- iucn[[i]][["name"]] 
}

## Reduce the lists down
iucn <- Reduce(rbind, iucn) 

# remove all of the extra species names in the list
iucn[1:341] <- NULL 

# Remove all lists with species with no threats
iucn <- iucn[sapply(iucn, length)>1] 

# binds all list elements by row to make a dataframe
iucn <- list.rbind(iucn)

# Delete the columns that will not be used in further analyses
iucn <- iucn[- c(3:7)] 

```


### Subset species at risk to threats

Here we subset all the seabird species that are known to be threatened from oil drilling, shipping lanes and subsistence fishing. 

IUCN Codes = https://www.iucnredlist.org/resources/threat-classification-scheme

-	3.1 Oil & gas drilling
-	4.3 Shipping lanes
-	5.4.1 Intentional use: subsistence/small scale (species being assessed is the target)[harvest]
-	5.4.3 Unintentional effects: subsistence/small scale (species being assessed is not the target)[harvest]


```{r, eval = FALSE}

# Subset based on the IUCN codes
Threats <- iucn[iucn$code %in% c("3.1", # Oil & gas drilling
                                 "4.3", # Shipping lanes
                                 "5.4.1", # Intentional subsistence fishing
                                 "5.4.3"), ] # Unintentional subsistence fishing

```


